---
title: "GY7702 Report"
author: "Author"
date: "23/12/2020"
output: pdf_document
---
This document was created to meet the requirements of GY7702 R for Data 
Science at University of Leicester. It was designed and created in R Markdown, a 
markup language that allows users to create documents that can be formatted to embed code blocks, code outputs and hyperlinks.
When the R Markdown file is compiled, the markup language is hidden and the document is displayed in plain text.

This content was created using [R](https://www.r-project.org/), [Rstudio](https://rstudio.com/), [RMarkdown](https://rmarkdown.rstudio.com/) and [GitHub](https://github.com/)


The libraries used in this assignment were 
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(gridExtra)
library(psych)
library(magrittr)
```

For further information regarding the source code, data and libraries used in this assignment, please see the projects GitHub page (). 

## References
The Author would like to acknowledge that this document includes teaching materials from Dr Stefano De Sabbato for the module 
[GY7702 R for Data Science](https://le.ac.uk/modules/2020/gy7702). Dr Stefanos teaching materials can be found [here](https://sdesabbata.github.io/granolarr/)  

R for Data Science by Garrett Grolemund and Hadley Wickham, O’Reilly Media, 2016. [See online book](https://r4ds.had.co.nz/)

# Load in the data 

```{r, message=FALSE, warning = FALSE}

# Extract 
unzip("Data/GY7702_2020-21_Assignment_2--data_pack.zip", exdir = "Data")

# Load
OAC_2011 <- 
  readr::read_csv(
    "Data/GY7702_2020-21_Assignment_2--data_pack/2011_OAC_Raw_kVariables.csv")

OAC_2011_meta <- 
  readr::read_csv(
    "Data/GY7702_2020-21_Assignment_2--data_pack/OA11_LSOA11_MSOA11_LAD11_EW_LUv2.csv")


```


# Option A 

## Joining the data sets

```{r, message = FALSE}
# Rename the column 'OA11CD' to OA so it matches the same column in OAC_2011 
OAC_2011_meta <- OAC_2011_meta %>% 
  dplyr::rename(
    OA = OA11CD
  )

# Join OAC_2011 with OAC_2011_meta, based on the OA column for each data frame 
complete_OAC <- OAC_2011 %>%
  dplyr::left_join(
    .,OAC_2011_meta
  )
```


## Subsetting 
* Selecting assigned LAD
* Selecting the appropriate variables
  + k004 Persons aged 45 to 64
  + k009 Persons aged over 16 who are single
  + k010 Persons aged over 16 who are married or in a registered same-sex civil partnership
  + k027 Households who live in a detached house or bungalow
  + k031 Households who own or have shared ownership of property
  + k041 Households with two or more cars or vans
  + k046 Employed persons aged between 16 and 74 who work part-time

```{r}
# Subset data based off LAD11NM = Wellingborough
wellingborough <- complete_OAC %>%
  dplyr::filter(
    LAD11NM == "Wellingborough"
    )

# Select the appropriate variables, based on their 'VariableCode' 
wellingborough_variables <- wellingborough %>%
  dplyr::select(
    k004, k009, k010, k027, k031, k041, k046
    )
```


# Question A.1 
Word count (excluding titles = 497)

## Exploring the distribution of Employed persons aged between 16 and 74 who work part-time
The distribution of part-time employed persons (PTEP) throughout 249 output areas (OA's) in the Local Area District (LAD) of Wellingborough was investigated.

### Basic histogram for visual interpretation 
```{r fig.align="center"}
k046_hist <- wellingborough %>%
  ggplot2::ggplot(aes(
    x = k046))+
  ggplot2::geom_bar()+
  scale_x_binned(
    n.breaks = 5)+
  xlab("Employed persons (16-74 years old) working part-time")+
  ylab("Frequency")+
  geom_vline(aes(
    xintercept = mean(k046)), 
    colour="red", 
    lwd = 1.5)+
  theme_gray(
    base_size = 15)

print(k046_hist)
```

An initial assessment suggests that the distribution of PTEPs throughout Wellingborough is normal. The majority of OA's in Welingborough have 35-45 PTEPS with some having as little as 20 and at least one having >80 PTEPs. A red line also demonstrates the average number of PTEPs per OA (38)



### Plotting against normal distribution
The distribution can be further assessed using ```stat_qq``` . This demonstrates where values may lie in the observed data that are affecting the normal distribution. 
```{r}
wellingborough %>%
  ggplot2::ggplot(aes(
    sample = k046))+
  ggplot2::stat_qq()+
  ggplot2::stat_qq_line(
    aes(
      colour = "normal distribution"), 
    show.legend = TRUE, 
    lwd = 1)+
  xlab("Theoretical normal distribution")+
  ylab("Observed value")+
  theme_gray(
    base_size = 15) 

```


There are an elevated number of PTEPs at the tails of the distribution. That is, a considerable number of OA's have a minimum and maximum number of PTEPs.    

### Shapiro-Wilk test 
Using a Shapiro-Wilk test we can investigate further if this a normal distribution. 
```{r}
k046 <- wellingborough %>%
  dplyr::pull(k046)%>%
stats::shapiro.test()
print(k046)
```

In the Shapiro-Wilk test, a null hypothesis that the data is normally distributed is set. This is rejected if the p-value <0.01. That is, there is a <1% chance of the data being normally distributed. In this case p < 0.01, so the data is not normally distributed. 

### Exploring skewness and kurtosis 
The overall shape of the data can be assessed further using ***skewness*** and ***kurtosis***

```{r}
wellingborough_skew <- 
  wellingborough %>%
    dplyr::select(k046)%>%
    pastecs::stat.desc(basic = FALSE, 
                       desc = FALSE, 
                       norm = TRUE
                       )
```
```{r, echo = FALSE}
kable(wellingborough_skew, format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped", "scale_up"))
```

A positive skew means the data leans left. The standard error for the skew ```skew.2SE``` lies outside of the range -1 + 1, meaning the result is significant. A positive kurtosis also means the data has heavy tails. However, this result is not significant as ```kurt.2SE``` lies within -1 +1



## Multi-variate analysis 
We have just carried out a uni-variate analysis. Relationships between variables can be assessed through multivariate analysis. 
The variables used in this study are split between collecting data on ***persons*** and ***households***. We can use a ```pairs.panel``` to compare these sets of data and discover correlations. 
The data here is not normally distributed (except k010), hence we will use ```kendall_tau``` which assumes a non-normal distribution and ties. 

```{r}
wellingborough_norm <- 
  wellingborough_variables %>%
  dplyr::select(
    k004, k009, k027, k031, k041, k046, k010
    ) %>%
  pastecs::stat.desc(
    norm = TRUE
    ) 

```

```{r, echo = FALSE}
kable(wellingborough_norm[19:20, ], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped", "scale_up"))
```

```{r}
wellingborough_variables %>%
  select(
    k004, k009, k027, k031, k041, k046, k010
    )%>%
  pairs.panels(
    method = "kendall", 
    stars = TRUE
    )
```


k010 has a strong correlation with k031 and k041. This is understandable. People who have shared ownership of a property will usually (but not always) be in a relationship and may also own a car. k009 does not positively correlate with any other variable, possibly for the opposite reasons.


### Normalising data 
The amount of people in a relationship (k010) and the amount of households with shared ownership (k031) is highly correlated (0.73). However, k031 gives information on ***households*** while k010 gives information on ***persons***. To aid comparison of the data sets, the z-scores can be taken.
```{r}
wellingborough_variables <- 
  wellingborough_variables %>%
    mutate(
      z_k031 = scale(k031)
      ) %>%
    mutate(
      z_k010 = scale(k010)
      )
```

Note, though that z scores conserve the distribution of the data sets. This means that a Shapiro-Wilks test will still fail for k031

```{r, echo = FALSE}
z_k031 <- wellingborough_variables %>%
  dplyr::pull(z_k031)%>%
stats::shapiro.test()
print(z_k031)
```

This assessment should not usually be carried out on non-normally distributed data. However, for the purpose of this study the new plot of z-scores offers more exploratory analysis. Values on either axis that are close to 0 are close to the mean. OA's that have the average number of people >16 in a relationship, also have the average number of households with shared ownership. 


```{r}
ggplot2::ggplot(
  wellingborough_variables, 
    aes(z_k031, 
        z_k010)
  )+
  ggplot2::geom_point()+
  xlab("households with shared ownership")+
  ylab("Persons >16 in a relationship")
```


# Question A.2
Word count (excluding titles = 480)
## Multiple linear regression

Households with two or more cars or vans (K041) and persons >16 who are married or in a same-sex civil partnership (K010) has a strong correlation with shared property ownership. Due to that, these variables will be selected to predict shared property ownership (k031). 

```{r, render = normal_print}
wellingborough_variables %$%
stats::lm(k031 ~ k010 + k041) -> 
  shared_prop_model

summary(shared_prop_model)
```


With a p <0.01, the performance of the model is good. A high F-statistic [f(2, 246)=543.3] demonstrates this model is more effective than if the mean for the outcome variable was used in all cases of the independent variables. An adjusted R^2 of 0.81 demonstrates that k010 and k041 explain 81% of the variance in k031.

For 1 increase in k010, k031 increases by 0.61. For 1 increase in k010, k031 increases by 0.17. This suggests that k010 has a larger impact on k031. The 3-dimensional model plane intersects the y-axis at 5.51. 

P-values show good results for k010 (P<0.01), however confidence is exceeded for k041 and the y-intercept which suggests the model is faltering.

### Standardise coefficients
```lm.beta``` standardises the coefficients to aid comparison. 1 standard deviation (SD) increment in k010, leads to 0.79 more shared properties and 1 SD increment in (k041) leads to 0.12 more shared properties. This further suggests that k010 is explaining more of the variance than k041. 


```{r}
shared_prop_model %>%
  lm.beta::lm.beta()
```
### Determine confidence intervals
```stats::confint()``` quantifies a range at which the true coefficients would occupy. If the range is small, the model is robust.

```{r}
shared_prop_model%>%
  stats::confint()
```
These ranges are relatively large. The model cannot be confident on the true value of the coefficients. The range is significantly smaller for k041 though. 

### Testing for outliers
We can test for outliers by assessing the standard residuals and the cook’s distance. A threshold for a cook’s distance > 1 will be set.  

```{r}

wellingborough_variables %>%
  mutate(
    # Take the standardised residuals, from the shared prop model 
    model_stdres = shared_prop_model %>% 
      stats::rstandard(),
    # Take the cook distance from the shared prop model 
    model_cook_dist = shared_prop_model %>% 
      stats::cooks.distance()
  ) -> 
  # Put those into a new table called shared_prop_output
  shared_prop_output

# From shared_prop_output, select the outcome variable with the associate 
# residuals and cook distance 
shared_prop_output %>%
  dplyr::select(
    k031, model_stdres, model_cook_dist
    ) %>%
  # Filter by stdres more than 2.58 and cook distance more than 1 
  dplyr::filter(
    abs(model_stdres) >2.58 | model_cook_dist >1
    )
```

```:filter(abs(model_stdres) >2.58``` is specifying any values above the 99% range of the data set. 2 values exceed this threshold. These values are outliers as they lie on the .5% upper and .5% lower limits of the data set.

### Testing assumptions 
We wil now step through the assumptions made during the building of this model 

1. **A linear relationship** \newline 
  This was previously demonstrated in the univariate analysis 

2. **Normal Distribution of standard residuals**  \newline 
A p>0.01 states that the distribution of the standard residual is not significant, and therefore normal.
```{r echo = FALSE}
shared_prop_output %$%
  stats::shapiro.test(model_stdres)
```

This can be visually assessed with a histogram
```{r fig.align="center", echo = FALSE}
shared_prop_output %>%
  ggplot2::ggplot(aes(
    x = model_stdres))+
  ggplot2::geom_bar()+
  scale_x_binned(
    n.breaks = 5)+
  xlab("Standard residuals")+
  ylab("Frequency")+
  theme_gray(
    base_size = 15)
```

3. **Standard residuals are homoscedastic and independent** \newline 
The homoscedasticity can be assessed by plotting the fitted values with the residuals 
```{r message=FALSE}
ggplot2::ggplot(shared_prop_model, 
                aes(
                  x = fitted.values(shared_prop_model), 
                  y = residuals(shared_prop_model)))+
  ggplot2::geom_point()+
  ggplot2::geom_smooth(se=FALSE)+
  xlab("Fitted values")+
  ylab("Residuals")+
  theme_gray(
    base_size = 15)
```

A "cloud" of residuals shows that residuals are not correlated with the model. Residuals are similar whether the outcome variable is high or low. However, a curved line demonstrates that these residuals can vary throughout the model.  

A Breusch-Pagan test further demonstrates that the residuals are homoscedastic as the result is not significant.  
```{r}
shared_prop_model %>%
  lmtest::bptest()
```


Independence of residuals can be tested with a Durbin-Watson test. The residuals are independent. 
```{r}
shared_prop_model %>%
  lmtest::dwtest()
```

4. **Relationship between independent variables is independent** \newline 
Finally, we can test the variance inflation factor (ViF). This tests if the variables are independent from one another. 
```{r}
shared_prop_model %>%
  car::vif()
```

The results for this test are relatively high. This means that the variables 'move' together. A change in k041 may also affect k010. These variables are not independent from one another.

